---
title: "BentoBox Meta Functions"
output:
    rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{BentoBox Meta Functions}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    fig.align = "center",
    fig.width = 3,
    fig.height = 3,
    collapse = TRUE,
    comment = "#>",
    warning = FALSE,
    message = FALSE
)
library(grid)
library(BentoBox)
library(BentoBoxData)
```

`BentoBox` meta functions enhance the `BentoBox` user experience by providing 
simple methods to display various genomic assembly data, simplify `BentoBox`
code, and construct `BentoBox` objects. Functions in this category include:

- **`bbGenomes`** and **`bbDefaultPackages`**

Simple functions to display the strings of available default genomic builds and 
the genomic annotation packages associated with these builds in `BentoBox` functions.

```{r}
bbGenomes()
```
```{r}
bbDefaultPackages("hg19")
```

- **`bbAssembly`**

A constructor to make custom combinations of genomic annotation packages for 
use in `BentoBox` functions through the `assembly` parameter.

- **`bbParams`**

A constructor to capture sets of parameters to be shared across multiple
function calls. `bbParams` objects can hold any argument from any `BentoBox`
function. Most often, `bbParams` objects are used to store a common genomic
region and common x-placement coordinate information. For a detailed example 
using the `bbParams` object, refer to the vignette 
[Plotting Multi-omic Data](plotting_multiomic_data.html#using-the-bbparams-object-1).

- **`colorby`** and **`bbMapColors`**

The `colorby` constructor allows us to color the data elements in `BentoBox` 
plots by various data features. These features can be a numerical range, like 
some kind of score value, or categorical values, like positive or negative 
strand. The `colorby` object is constructed by specifying the name of the data 
column to color by, an optional color palette function, and an optional range 
for numerical values. If not specified, `BentoBox` will use the `RColorBrewer`
"YlGnBl" palette for mapping numerical data and the "Pairs" palette for 
qualitative data.

For example, if we revist the BED plot above, `IMR90_ChIP_CTCF_reads` has an additional 
`strand` column for each BED element: 

```{r eval = TRUE, echo = TRUE}
data("IMR90_ChIP_CTCF_reads")
head(IMR90_ChIP_CTCF_reads)
```

Thus, we can use the `colorby` constructor to color BED elements by positive 
or negative strand. The `strand` column will be converted to a factor with a 
`-` level and `+` level. These values will be mapped to our input `palette`:

```{r eval=FALSE, echo=TRUE}
set.seed(nrow(IMR90_ChIP_CTCF_reads))
bbPlotRanges(
    data = IMR90_ChIP_CTCF_reads,
    chrom = "chr21", chromstart = 29073000, chromend = 29074000,
    assembly = "hg19",
    fill = colorby("strand", palette = 
                       colorRampPalette(c("#7ecdbb", "#37a7db"))),
    x = 0.5, y = 0.25, width = 6.5, height = 4.25,
    just = c("left", "top"), default.units = "inches"
)
```

```{r aes_03, echo=FALSE, fig.height=6, fig.width=9.5}

bbPageCreate(
    width = 7.5, height = 5, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)

set.seed(nrow(IMR90_ChIP_CTCF_reads))
bbPlotRanges(
    data = IMR90_ChIP_CTCF_reads,
    chrom = "chr21", chromstart = 29073000, chromend = 29074000,
    assembly = "hg19",
    fill = colorby("strand", palette = 
                       colorRampPalette(c("#7ecdbb", "#37a7db"))),
    strandSplit = TRUE,
    x = 0.5, y = 0.25, width = 6.5, height = 4.25,
    just = c("left", "top"), default.units = "inches"
)
```

To further control the order of color mapping, we can set our categorical 
`colorby` column as a factor with our own order of levels before plotting:

```{r eval=TRUE, echo=TRUE}
data("IMR90_ChIP_CTCF_reads")
IMR90_ChIP_CTCF_reads$strand <- factor(IMR90_ChIP_CTCF_reads$strand, levels = c("+", "-"))
head(IMR90_ChIP_CTCF_reads$strand)
```

Now we've set the `+` level as our first level, so our `palette` will map
colors in the opposite order from before:

```{r aes_04, echo=FALSE, fig.height=6, fig.width=9.5}

bbPageCreate(
    width = 7.5, height = 5, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
set.seed(nrow(IMR90_ChIP_CTCF_reads))
bbPlotRanges(
    data = IMR90_ChIP_CTCF_reads,
    chrom = "chr21", chromstart = 29073000, chromend = 29074000,
    assembly = "hg19",
    fill = colorby("strand", palette = 
                       colorRampPalette(c("#7ecdbb", "#37a7db"))),
    strandSplit = TRUE,
    x = 0.5, y = 0.25, width = 6.5, height = 4.25,
    just = c("left", "top"), default.units = "inches"
)
```

In this example, we will color BEDPE arches by a range of numerical values 
we will add as a `length` column:

```{r eval=TRUE, echo=TRUE }
data("IMR90_DNAloops_pairs")
IMR90_DNAloops_pairs$length <- (IMR90_DNAloops_pairs$start2 - IMR90_DNAloops_pairs$start1) / 1000
head(IMR90_DNAloops_pairs$length)
```

Now we can set `fill` as a `colorby` object to color the BEDPE `length` column by:
```{r eval=FALSE, echo=TRUE}
bedpePlot <- bbPlotPairsArches(
    data = IMR90_DNAloops_pairs,
    chrom = "chr21", chromstart = 27900000, chromend = 30700000,
    assembly = "hg19",
    fill = colorby("length", 
                   palette = colorRampPalette(c("dodgerblue2", "firebrick2"))),
    linecolor = "fill",
    archHeight = IMR90_DNAloops_pairs$length / max(IMR90_DNAloops_pairs$length),
    alpha = 1,
    x = 0.25, y = 0.25, width = 7, height = 1.5,
    just = c("left", "top"),
    default.units = "inches"
)
```

```{r aes_05, echo=FALSE, fig.width=9.5, message=FALSE, height=3}

bbPageCreate(
    width = 7.5, height = 2, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
heights <- IMR90_DNAloops_pairs$length / max(IMR90_DNAloops_pairs$length)
bedpePlot <- bbPlotPairsArches(
    data = IMR90_DNAloops_pairs,
    chrom = "chr21", chromstart = 27900000, chromend = 30700000,
    assembly = "hg19",
    fill = colorby("length", 
                   palette = colorRampPalette(c("dodgerblue2", "firebrick2"))),
    linecolor = "fill",
    archHeight = heights, alpha = 1,
    x = 0.25, y = 0.25, width = 7, height = 1.5,
    just = c("left", "top"),
    default.units = "inches"
)
```

And now since we have numbers mapped to colors, we can use 
`bbAnnoHeatmapLegend()` with our `bb_arches` object to add a legend for 
the `colorby` we performed:

```{r eval=FALSE, echo=TRUE}
bbAnnoHeatmapLegend(
    plot = bedpePlot, fontcolor = "black",
    x = 7.0, y = 0.25,
    width = 0.10, height = 1, fontsize = 10
)
```
```{r aes_06, echo=FALSE, fig.width=9.5, message=FALSE, height=3}

bbPageCreate(
    width = 7.5, height = 2, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
heights <- IMR90_DNAloops_pairs$length / max(IMR90_DNAloops_pairs$length)
bedpePlot <- bbPlotPairsArches(
    data = IMR90_DNAloops_pairs,
    chrom = "chr21", chromstart = 27900000, chromend = 30700000,
    assembly = "hg19",
    fill = colorby("length", 
                   palette = colorRampPalette(c("dodgerblue2", "firebrick2"))),
    linecolor = "fill",
    archHeight = heights, alpha = 1,
    x = 0.25, y = 0.25, width = 7, height = 1.5,
    just = c("left", "top"),
    default.units = "inches"
)
bbAnnoHeatmapLegend(
    plot = bedpePlot, fontcolor = "black",
    x = 7.0, y = 0.25,
    width = 0.10, height = 1, fontsize = 10
)
```

If users wish to map values to a color palette before passing them into
a `BentoBox` function, they can use `bbMapColors`:
```{r echo=TRUE, eval=FALSE}
colors <- bbMapColors(vector = IMR90_DNAloops_pairs$length,
             palette = colorRampPalette(c("dodgerblue2", "firebrick2")))

bedpePlot <- bbPlotPairsArches(
    data = IMR90_DNAloops_pairs,
    chrom = "chr21", chromstart = 27900000, chromend = 30700000,
    assembly = "hg19",
    fill = colors,
    linecolor = "fill",
    archHeight = heights, alpha = 1,
    x = 0.25, y = 0.25, width = 7, height = 1.5,
    just = c("left", "top"),
    default.units = "inches"
)

```

```{r aes_07, echo=FALSE, fig.width=9.5, message=FALSE, height=3}

bbPageCreate(
    width = 7.5, height = 2, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
heights <- IMR90_DNAloops_pairs$length / max(IMR90_DNAloops_pairs$length)
colors <- bbMapColors(vector = IMR90_DNAloops_pairs$length,
             palette = colorRampPalette(c("dodgerblue2", "firebrick2")))
bedpePlot <- bbPlotPairsArches(
    data = IMR90_DNAloops_pairs,
    chrom = "chr21", chromstart = 27900000, chromend = 30700000,
    assembly = "hg19",
    fill = colors,
    linecolor = "fill",
    archHeight = heights, alpha = 1,
    x = 0.25, y = 0.25, width = 7, height = 1.5,
    just = c("left", "top"),
    default.units = "inches"
)
bbAnnoHeatmapLegend(
    plot = bedpePlot, fontcolor = "black",
    x = 7.0, y = 0.25,
    width = 0.10, height = 1, fontsize = 10
)
```

### Session Info
```{r}
sessionInfo()
```