---
title: The BentoBox Page
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{The BentoBox Page}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    message = FALSE,
    warning = FALSE
)
library(BentoBox)
library(grid)
library(BentoBoxData)
```

`BentoBox` uses a coordinate-based plotting system to define the size and 
location of plots. This system makes the plotting process intuitive and 
absolute, meaning that plots cannot be squished and stretched based 
on their *relative* sizes. This also allows for precise control of the size 
of each visualization and the location of all plots, annotations, and text. 

All `BentoBox` page functions begin with the `bbPage` prefix. 

Users can create a page in their preferred size and unit of measurement using
`bbPageCreate()`. Within this function the user can also set gridlines in 
the vertical and horizontal directions with `xgrid` and `ygrid`, respectively. 
By default these values are set to 0.5 of the unit. In the following example 
we demonstrate creating a standard 8.5 x 11 inch page:

```{r page_demo01, fig.width=11, fig.height=12}
bbPageCreate(width = 8.5, height = 11, default.units = "inches")
```

Or we could create a smaller sized page in a different set of units 
with different gridlines:

```{r page_demo02, fig.width=5, fig.height=4}
bbPageCreate(width = 8, height = 8, xgrid = 1, ygrid = 1, default.units = "cm")
```

We could turn off gridlines entirely by setting `xgrid` and `ygrid` to 0:

```{r page_demo03, fig.width=5, fig.height=4}
bbPageCreate(
    width = 3, height = 3, xgrid = 0, ygrid = 0,
    default.units = "inches"
)
```

If we want more specific gridlines on our page, we can use 
the `bbPageGuideHorizontal()` and `bbPageGuideVertical()` functions:

```{r page_demo04, fig.width=5, fig.height=4}
bbPageCreate(width = 3, height = 3, default.units = "inches")
## Add a horizontal guide at y = 2.25 inches
bbPageGuideHorizontal(y = 2.25, default.units = "inches")
## Add a vertical guide at x = 0.75 inches
bbPageGuideVertical(x = 0.75, default.units = "inches")
```

We can also remove all guidelines from the plot once we are finished using 
guides by using the `bbPageGuideHide()` function:

```{r page_demo05, fig.width=5, fig.height=4}
## Create page
bbPageCreate(width = 3, height = 3, default.units = "inches")
## Remove guides
bbPageGuideHide()
```

## Coordinate systems and units

`BentoBox` is compatible with numerous coordinate systems, which are 
flexible enough to be used in combination. Brief descriptions of the most 
commonly used `BentoBox` coordinate systems are as follows:

+-------------------+---------------------------------------------------------+
| Coordinate System | Description                                             |
+==================:+:========================================================+
| **"npc"**         | Normalized Parent Coordinates. Treats the bottom-left   |
|                   | corner of the plotting region as the location (0,0) and |
|                   | the top-right corner as (1,1).                          |
+-------------------+---------------------------------------------------------+
| **"snpc"**        | Squared Normalized Parent Coordiantes. Placements and   |
|                   | sizes are expressed as a proportion of the smaller of   |
|                   | the width and height of the plotting region.            |
+-------------------+---------------------------------------------------------+
| **"native"**      | Placements and sizes are relative to the x- and         |
|                   | y-scales of the plotting region.                        |
+-------------------+---------------------------------------------------------+
| **"inches"**      | Placements and sizes are in terms of physical inches.   |
+-------------------+---------------------------------------------------------+
| **"cm"**          | Placements and sizes are in terms of physical           |
|                   | centimeters.                                            |
+-------------------+---------------------------------------------------------+
| **"mm"**          | Placements and sizes are in terms of physical           |
|                   | millimeters.                                            |
+-------------------+---------------------------------------------------------+
| **"points"**      | Placements and sizes are in terms of physical points.   |
|                   | There are 72.27 points per inch.                        |
+-------------------+---------------------------------------------------------+

We can set the `bbPage` in one coordinate system, but then place and arrange 
our plots using other coordinate systems. For example, we can set our page 
as 3 x 3 **inches**:

```{r coord_01, eval=TRUE, message=FALSE,echo=TRUE, fig.width=5, fig.height=4}
bbPageCreate(width = 3, height = 3, default.units = "inches")
```

But we can then switch to **npc** coordinates to plot something in the 
center of the page at (0.5, 0.5) npc. The `unit()` function allows us to 
easily specify `x`, `y`, `width`, and `height` in combinations of different 
units in one plotting function call. 
```{r eval=FALSE, message=FALSE,echo=TRUE}
bbPlotRect(
    x = unit(0.5, "npc"), y = unit(0.5, "npc"), width = 1, height = 1,
    default.units = "inches"
)
```
```{r coord_02, eval=TRUE, message=FALSE,echo=FALSE, fig.width=5, fig.height=4}
bbPageCreate(width = 3, height = 3, default.units = "inches")
bbPlotRect(x = unit(0.5, "npc"), y = unit(0.5, "npc"), width = 1, height = 1)
```

The **native** coordinate system is particularly useful for annotation 
functions that can plot relative to the genomic scales of a plot. For example,
we can use `bbAnnoText()` to annotate some text at a specific genomic 
location in a plot:

```{r coord_03, echo=TRUE, fig.height=2.5, fig.width=7, message=FALSE}
bbPageCreate(
    width = 5, height = 1.5, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)

library(BentoBoxData)
data("IMR90_ChIP_H3K27ac_signal")
signalPlot <- bbPlotSignal(
    data = IMR90_ChIP_H3K27ac_signal,
    chrom = "chr21", chromstart = 28000000, chromend = 30300000,
    assembly = "hg19",
    x = 0.5, y = 0.25, width = 4, height = 0.75, default.units = "inches"
)
bbAnnoGenomeLabel(plot = signalPlot, x = 0.5, y = 1.01)

## Annotate text at average x-coordinate of data peak
peakScore <- IMR90_ChIP_H3K27ac_signal[which(
    IMR90_ChIP_H3K27ac_signal$score == max(IMR90_ChIP_H3K27ac_signal$score)
), ]
peakPos <- round((min(peakScore$start) + max(peakScore$end)) * 0.5)
bbAnnoText(
    plot = signalPlot, label = format(peakPos, big.mark = ","), fontsize = 8,
    x = unit(peakPos, "native"), y = unit(1, "npc"),
    just = "bottom"
)
```

## Working with plot objects

In `BentoBox` all plot objects are boxes, with user-defined positions and 
sizes. All plot objects can be placed on a `bbPage` using the placement 
arguments (e.g. `x`, `y`, `width`, `height`, `just`, `default.units`, ...). 
The `bbPage` sets the origin of the plot at the top left corner of the page. 
By default, the `x` and `y` arguments place the top-left corner of a plot in 
the specified position on the `bbPage` while the `width` and `height` 
arguments define the size of the plot.

For example, if users want the top-left corner of their plot to be 0.5 inches 
down from the top of the page and 0.5 inches from the left...

```{r arrange_example01, echo=FALSE, fig.height=4, fig.width=5, message=FALSE}
## Create page
bbPageCreate(width = 3, height = 3, default.units = "inches")

## Draw arrows
bbPlotSegments(
    x0 = 0.5, y0 = 0, x1 = 0.5, y1 = 0.5,
    arrow = arrow(
        angle = 30,
        length = unit(0.1, "inches"),
        ends = "last",
        type = "closed"
    ),
    fill = "black"
)

bbPlotSegments(
    x0 = 0, y0 = 0.5, x1 = 0.5, y1 = 0.5,
    arrow = arrow(
        angle = 30,
        length = unit(0.1, "inches"),
        ends = "last",
        type = "closed"
    ),
    fill = "black"
)

## Draw text
bbPlotText(label = "0.5 in", x = 0.6, y = 0.25, just = "left")
bbPlotText(label = "0.5 in", x = 0.25, y = 0.6, just = "top")
```

and the plot to be 2 inches wide and 1 inch tall...

```{r arrange_example02, echo=FALSE, fig.height=4, fig.width=5, message=FALSE}
## Create page
bbPageCreate(width = 3, height = 3, default.units = "inches")

## Draw arrows
bbPlotSegments(
    x0 = 0.5, y0 = 0, x1 = 0.5, y1 = 0.5,
    arrow = arrow(
        angle = 30,
        length = unit(0.1, "inches"),
        ends = "last",
        type = "closed"
    ),
    fill = "black"
)

bbPlotSegments(
    x0 = 0, y0 = 0.5, x1 = 0.5, y1 = 0.5,
    arrow = arrow(
        angle = 30,
        length = unit(0.1, "inches"),
        ends = "last",
        type = "closed"
    ),
    fill = "black"
)

## Draw plot dimension arrows
bbPlotSegments(
    x0 = 0.5, y0 = 0.5, x1 = 2.5, y1 = 0.5,
    arrow = arrow(
        angle = 30,
        length = unit(0.1, "inches"),
        ends = "both",
        type = "closed"
    ),
    fill = "black"
)

bbPlotSegments(
    x0 = 0.5, y0 = 0.5, x1 = 0.5, y1 = 1.5,
    arrow = arrow(
        angle = 30,
        length = unit(0.1, "inches"),
        ends = "both",
        type = "closed"
    ),
    fill = "black"
)

## Draw text
bbPlotText(label = "2 in", x = 1.5, y = 0.4, just = "bottom")
bbPlotText(label = "1 in", x = 0.4, y = 1, just = "right")
```

`BentoBox` can make the plot with these exact dimensions:

```{r arrange_example03, echo=TRUE, fig.height=4, fig.width=5, message=FALSE}
## Create page
bbPageCreate(width = 3, height = 3, default.units = "inches")

## Plot rectangle
bbPlotRect(
    x = 0.5, y = 0.5, width = 2, height = 1,
    just = c("left", "top"), default.units = "inches"
)
```

`BentoBox` also provides the helper function `bbPagePlotPlace()` for placing 
plot objects that have been previously defined:

```{r arrange_example04, echo=TRUE, fig.height=4, fig.width=5, message=FALSE}
## Load data
library(BentoBoxData)
data("IMR90_ChIP_H3K27ac_signal")

## Create page
bbPageCreate(width = 3, height = 3, default.units = "inches")

## Define signal plot
signalPlot <- bbPlotSignal(
    data = IMR90_ChIP_H3K27ac_signal,
    chrom = "chr21", chromstart = 28000000, chromend = 30300000,
    assembly = "hg19",
    draw = FALSE
)

## Place plot on bbPage
bbPagePlotPlace(
    plot = signalPlot,
    x = 0.5, y = 0.5, width = 2, height = 1,
    just = c("left", "top"), default.units = "inches"
)
```

and `bbPagePlotRemove()` for removing plots from a page:

```{r arrange_example05, echo=TRUE, fig.height=4, fig.width=5, message=FALSE}
# Load data
library(BentoBoxData)
data("IMR90_ChIP_H3K27ac_signal")

## Create page
bbPageCreate(width = 3, height = 3, default.units = "inches")

## Plot and place signal plot
signalPlot <- bbPlotSignal(
    data = IMR90_ChIP_H3K27ac_signal,
    chrom = "chr21", chromstart = 28000000, chromend = 30300000,
    assembly = "hg19",
    x = 0.5, y = 0.5, width = 2, height = 1,
    just = c("left", "top"), default.units = "inches"
)
## Remove signal plot
bbPagePlotRemove(plot = signalPlot)
```


These functions give the users additional flexibility in how they create 
their R scripts and `BentoBox` layouts.

## Using the `just` parameter

While the `x`, `y`, `width`, and `height` parameters are relative to the 
top-left corner of the plot by default, the `just` parameter provides 
additional flexibility by allowing users to change the placement reference 
point. The `just` parameter accepts a character or numeric vector of length 2 
describing the horizontal and vertical justification (or reference point), 
respectively.

The just parameter can be set using character strings `"left"`, `"right"`, 
`"center"`, `"bottom"` and `"top"`:

```{r arrange_example06, echo=FALSE, fig.height=5, fig.width=6, message=FALSE}
## Load libraries
library(BentoBox)

## Create page
bbPageCreate(
    width = 4, height = 4, xgrid = 1, ygrid = 1,
    default.units = "inches"
)

## Define x and y coords, colors, and justifications
xcoords <- c(0.25, 2, 3.75, 3.75, 3.75, 2, 0.25, 0.25, 2)
ycoords <- c(0.25, 0.25, 0.25, 2, 3.75, 3.75, 3.75, 2, 2)
cols <- c(RColorBrewer::brewer.pal(n = 8, name = "Dark2"), "black")
justs <- list(
    c("left", "top"), c("center", "top"), c("right", "top"),
    c("right", "center"), c("right", "bottom"), c("center", "bottom"),
    c("left", "bottom"), c("left", "center"), c("center", "center")
)

## Define x and y shift vectors for adjusting text
s <- 0.1
xshift <- c(s, 0, -s, -s, -s, 0, s, s, 0)
yshift <- c(s, s, s, 0, -s, -s, -s, 0, -s * 3)

## Draw guides
bbPageGuideHorizontal(y = ycoords, linecolor = "grey90", lty = 2)
bbPageGuideVertical(x = xcoords, linecolor = "grey90", lty = 2)

## Draw box
bbPlotRect(
    x = 0.25, y = 0.25, width = 3.5, height = 3.5,
    just = c("left", "top")
)

## Plot points
bbPlotCircle(
    x = xcoords,
    y = ycoords,
    r = 0.035, linecolor = cols, fill = cols
)

## Text descriptions
for (i in 1:9) {
    bbPlotText(
        label = sprintf(
            "x = %s\ny = %s\njust = c('%s', '%s')",
            xcoords[i], ycoords[i], justs[[i]][1], justs[[i]][2]
        ),
        x = xcoords[i] + xshift[i],
        y = ycoords[i] + yshift[i],
        just = justs[[i]],
        fontsize = 6.5, fontcolor = cols[i], fontface = "bold"
    )
}
```

Or it can be set using numeric values where 0 means left/bottom, 1 means 
right/top, and 0.5 means center: 

```{r arrange_example07, echo=FALSE, fig.height=5, fig.width=6, message=FALSE}
## Load libraries
library(BentoBox)

## Create page
bbPageCreate(
    width = 4, height = 4, xgrid = 1, ygrid = 1,
    default.units = "inches"
)

## Define x and y coords, colors, and justifications
xcoords <- c(0.25, 2, 3.75, 3.75, 3.75, 2, 0.25, 0.25, 2)
ycoords <- c(0.25, 0.25, 0.25, 2, 3.75, 3.75, 3.75, 2, 2)
cols <- c(RColorBrewer::brewer.pal(n = 8, name = "Dark2"), "black")
justs <- list(
    c(0, 1), c(0.5, 1), c(1, 1),
    c(1, 0.5), c(1, 0), c(0.5, 0),
    c(0, 0), c(0, 0.5), c(0.5, 0.5)
)

## Define x and y shift vectors for adjusting text
s <- 0.1
xshift <- c(s, 0, -s, -s, -s, 0, s, s, 0)
yshift <- c(s, s, s, 0, -s, -s, -s, 0, -s * 3)

## Draw guides
bbPageGuideHorizontal(y = ycoords, linecolor = "grey90", lty = 2)
bbPageGuideVertical(x = xcoords, linecolor = "grey90", lty = 2)

## Draw box
bbPlotRect(
    x = 0.25, y = 0.25, width = 3.5, height = 3.5,
    just = c("left", "top")
)

## Plot points
bbPlotCircle(
    x = xcoords,
    y = ycoords,
    r = 0.035, linecolor = cols, fill = cols
)

## Text descriptions
for (i in 1:9) {
    bbPlotText(
        label = sprintf(
            "x = %s\ny = %s\njust = c(%s, %s)",
            xcoords[i], ycoords[i], justs[[i]][1], justs[[i]][2]
        ),
        x = xcoords[i] + xshift[i],
        y = ycoords[i] + yshift[i],
        just = justs[[i]],
        fontsize = 6.5, fontcolor = cols[i], fontface = "bold"
    )
}
```

This is particularly useful when an object needs to be aligned in reference 
to another plot object or page marker. For example, in the Hi-C plot below we 
might want to align the top-right corner of the heatmap legend to the 3-inch 
mark. There is no need to calculate the top-left position 
(i.e. `3 inches - (legend width)`) to determine where to place the heatmap 
legend. Instead we can change the `just` parameter to `just=c('right', 'top')`:

```{r arrange_example08, echo=TRUE, fig.height=5, fig.width=6, message=FALSE}
## Load example Hi-C data
library(BentoBoxData)
data("IMR90_HiC_10kb")

## Create a BentoBox page
bbPageCreate(width = 3.25, height = 3.25, default.units = "inches")

## Plot Hi-C data with placing information
hicPlot <- bbPlotHicSquare(
    data = IMR90_HiC_10kb,
    chrom = "chr21", chromstart = 28000000, chromend = 30300000,
    assembly = "hg19",
    x = 0.25, y = 0.25, width = 2.5, height = 2.5, default.units = "inches"
)

## Add color scale annotation with just = c("right", "top")
bbAnnoHeatmapLegend(
    plot = hicPlot,
    x = 3, y = 0.25, width = 0.1, height = 1.25,
    just = c("right", "top"), default.units = "inches"
)
```

# Session Info
```{r}
sessionInfo()
```