---
title: "Incorporating `ggplots`"
output:
    rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Incorporating `ggplots`}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    fig.align = "center",
    fig.width = 3,
    fig.height = 3,
    collapse = TRUE,
    comment = "#>",
    warning = FALSE,
    message = FALSE
)
library(grid)
library(plotgardener)
library(plotgardenerData)
```

```{r eval=TRUE, echo=FALSE, message=FALSE}
library(showtext)
font_add(
    family = "ProximaNova",
    regular = system.file("extdata",
                        "proximanova-regular.otf",
                        package = "plotgardener"
    ),
    bold = system.file("extdata",
                    "proximanova-semibold.otf",
                    package = "plotgardener"
    ),
    italic = system.file("extdata",
                        "proximanova-regitalic.otf",
                        package = "plotgardener")
)
showtext_auto()
```

In addition to its numerous genomic functions, `plotgardener` can size and place 
`ggplots` within a `plotgardener` layout. Rather than arranging `ggplots` in a 
relative manner, `plotgardener` can make and place `ggplots` in absolute sizes 
and locations. This makes it simple and intuitive to make complex `ggplot` 
arrangements beyond a basic grid-style layout.

For example, let's say we wanted to make a complex multi-panel ggplot about 
COVID-19 data consisting of the following plots:

1) A United States map depicting COVID-19 cases:

```{r ggplot_01, echo=TRUE, fig.showtext=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
data("COVID_USA_cases")

US_map <- ggplot(COVID_USA_cases, aes(long, lat, group = group)) +
    theme_void() +
    geom_polygon(aes(fill = cases_100K), color = "white", size = 0.3) +
    scale_fill_distiller(
        palette = "YlGnBu", direction = 1,
        labels = label_number(suffix = "", scale = 1e-3, accuracy = 1)
    ) +
    theme(
        legend.position = "left",
        legend.justification = c(0.5, 0.95),
        legend.title = element_blank(),
        legend.text = element_text(size = 7),
        legend.key.width = unit(0.3, "cm"),
        legend.key.height = unit(0.4, "cm"),
        plot.title = element_text(
            hjust = 0, vjust = -1,
            family = "ProximaNova", face = "bold",
            size = 12
        ),
        plot.title.position = "plot"
    ) +
    labs(title = "Thousands of COVID-19 Cases per 100,000 People") +
    coord_fixed(1.3)

print(US_map)
```

2) Line plots showing the accumulation of COVID-19 cases over time:

```{r ggplot_02, echo=TRUE, fig.showtext=TRUE, message=FALSE}
data("COVID_NY_FL_tracking")

# Format y-labels
ylabels <- seq(0, 2000000, by = 500000) / 1e6
ylabels[c(3, 5)] <- round(ylabels[c(3, 5)], digits = 0)
ylabels[c(2, 4)] <- round(ylabels[c(2, 4)], digits = 1)
ylabels[5] <- paste0(ylabels[5], "M cases")
ylabels[1] <- ""

casesNY <- COVID_NY_FL_tracking[COVID_NY_FL_tracking$state == "new york", ]
casesNYpoly <- rbind(
    casesNY,
    data.frame(
        "date" = as.Date("2021-03-07"),
        "state" = "new york",
        "caseIncrease" = -1 * sum(casesNY$caseIncrease)
    )
)

cases_NYline <- ggplot(
    casesNY,
    aes(x = date, y = cumsum(caseIncrease))
) +
    geom_polygon(data = casesNYpoly, fill = "#B8E6E6") +
    scale_x_date(
        labels = date_format("%b '%y"),
        breaks = as.Date(c("2020-05-01", "2020-09-01", "2021-01-01")),
        limits = as.Date(c("2020-01-29", "2021-03-07")),
        expand = c(0, 0)
    ) +
    scale_y_continuous(labels = ylabels, position = "right", expand = c(0, 0)) +
    geom_hline(
        yintercept = c(500000, 1000000, 1500000, 2000000),
        color = "white", linetype = "dashed", size = 0.3
    ) +
    coord_cartesian(ylim = c(0, 2000000)) +
    theme(
        panel.background = element_rect(fill = "transparent", color = NA),
        text = element_text(family = "ProximaNova"),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "transparent", color = NA),
        axis.line.x.bottom = element_blank(),
        axis.line.y = element_line(size = 0.1, color = "#8F9BB3"),
        axis.text.x = element_text(
            size = 7, hjust = 0.5,
            margin = margin(t = -10), color = "black"
        ),
        axis.title.x = element_blank(),
        axis.ticks.x = element_line(size = 0.2, color = "black"),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 7, color = "black"),
        axis.ticks.y = element_blank(),
        axis.ticks.length.x.bottom = unit(-0.1, "cm"),
        plot.title = element_text(size = 8, hjust = 1),
        plot.title.position = "plot"
    )

print(cases_NYline)
```

```{r ggplot_03, echo=TRUE, fig.showtext=TRUE, message=FALSE}
casesFL <- COVID_NY_FL_tracking[COVID_NY_FL_tracking$state == "florida", ]
casesFLpoly <- rbind(
    casesFL,
    data.frame(
        "date" = as.Date("2021-03-07"),
        "state" = "florida",
        "caseIncrease" = -1 * sum(casesFL$caseIncrease)
    )
)

cases_FLline <- ggplot(
    casesFL,
    aes(x = date, y = cumsum(caseIncrease))
) +
    geom_polygon(data = casesFLpoly, fill = "#B8E6E6") +
    scale_x_date(
        labels = date_format("%b '%y"),
        breaks = as.Date(c("2020-05-01", "2020-09-01", "2021-01-01")),
        limits = as.Date(c("2020-01-29", "2021-03-07")),
        expand = c(0, 0)
    ) +
    scale_y_continuous(labels = ylabels, position = "right", expand = c(0, 0)) +
    geom_hline(
        yintercept = c(500000, 1000000, 1500000, 2000000),
        color = "white", linetype = "dashed", size = 0.3
    ) +
    coord_cartesian(ylim = c(0, 2000000)) +
    theme(
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA),
        text = element_text(family = "ProximaNova"),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line.x.bottom = element_blank(),
        axis.line.y = element_line(size = 0.1, color = "#8F9BB3"),
        axis.title = element_blank(),
        axis.text.y = element_text(size = 7, color = "black"),
        axis.text.x = element_text(
            size = 7, hjust = 0.5,
            margin = margin(t = -10), color = "black"
        ),
        axis.ticks = element_line(color = "black", size = 0.2),
        axis.ticks.y = element_blank(),
        axis.ticks.length.x.bottom = unit(-0.1, "cm"),
        plot.title = element_text(size = 8, hjust = 1),
        plot.title.position = "plot"
    )

print(cases_FLline)
```

3) Pie charts of COVID-19 vaccination status:

```{r ggplot_04, echo=TRUE, fig.showtext=TRUE, message=FALSE}
data("COVID_NY_FL_vaccines")

vaccines_NYpie <- ggplot(
    COVID_NY_FL_vaccines[COVID_NY_FL_vaccines$state == "new york", ],
    aes(x = "", y = value, fill = vax_group)
) +
    geom_bar(width = 1, stat = "identity") +
    theme_void() +
    scale_fill_manual(values = c("#FBAA7E", "#F7EEBF", "#FBCB88")) +
    coord_polar(theta = "y", start = 2.125, clip = "off") +
    geom_text(aes(
        x = c(1.9, 2, 1.9),
        y = c(1.65e7, 1.3e6, 7.8e6),
        label = paste0(percent, "%")
    ),
    size = 2.25, color = "black"
    ) +
    theme(
        legend.position = "none",
        plot.title = element_text(
            hjust = 0.5, vjust = -3.5, size = 10,
            family = "ProximaNova", face = "bold"
        ),
        text = element_text(family = "ProximaNova")
    ) +
    labs(title = "New York")

print(vaccines_NYpie)
```
```{r ggplot_05, echo = TRUE, eval = TRUE}
vaccines_FLpie <- ggplot(
    COVID_NY_FL_vaccines[COVID_NY_FL_vaccines$state == "florida", ],
    aes(x = "", y = value, fill = vax_group)
) +
    geom_bar(width = 1, stat = "identity") +
    scale_fill_manual(values = c("#FBAA7E", "#F7EEBF", "#FBCB88")) +
    theme_void() +
    coord_polar(theta = "y", start = pi / 1.78, clip = "off") +
    geom_text(aes(
        x = c(1.95, 2, 1.9),
        y = c(1.9e7, 1.83e6, 9.6e6),
        label = paste0(percent, "%")
    ),
    color = "black",
    size = 2.25
    ) +
    theme(
        legend.position = "none",
        plot.title = element_text(
            hjust = 0.5, vjust = -4, size = 10,
            family = "ProximaNova", face = "bold"
        ),
        text = element_text(family = "ProximaNova")
    ) +
    labs(title = "Florida")

print(vaccines_FLpie)
```

We can now easily overlap and size all these `ggplots` by passing our saved 
plot objects into `plotGG()`:

```{r gg_plot05, echo=TRUE, fig.height=4.5, fig.showtext=TRUE, fig.width=11.5}
pageCreate(width = 9.5, height = 3.5, default.units = "inches")

plotGG(
    plot = US_map,
    x = 0.1, y = 0,
    width = 6.5, height = 3.5, just = c("left", "top")
)
plotGG(
    plot = cases_NYline,
    x = 6.25, y = 1.8,
    width = 3.025, height = 1.4, just = c("left", "bottom")
)
plotGG(
    plot = cases_FLline,
    x = 6.25, y = 3.5,
    width = 3.025, height = 1.4, just = c("left", "bottom")
)
```

In particular, `plotgardener` makes it easy to resize and place our pie charts in
a layout that overlaps our line plots without it affecting the sizing of the 
other plots on the page:

```{r eval = FALSE, echo = TRUE}
plotGG(
    plot = vaccines_NYpie,
    x = 6.37, y = -0.05,
    width = 1.45, height = 1.45, just = c("left", "top")
)
plotGG(
    plot = vaccines_FLpie,
    x = 6.37, y = 1.67,
    width = 1.45, height = 1.45, just = c("left", "top")
)
```

```{r gg_plot06, echo=FALSE, fig.height=4.5, fig.showtext=TRUE, ,fig.width=11.5}
pageCreate(width = 9.5, height = 3.5, default.units = "inches")

plotGG(
    plot = US_map,
    x = 0.1, y = 0,
    width = 6.5, height = 3.5, just = c("left", "top")
)
plotGG(
    plot = cases_NYline,
    x = 6.25, y = 1.8,
    width = 3.025, height = 1.4, just = c("left", "bottom")
)
plotGG(
    plot = cases_FLline,
    x = 6.25, y = 3.5,
    width = 3.025, height = 1.4, just = c("left", "bottom")
)
plotGG(
    plot = vaccines_NYpie,
    x = 6.37, y = -0.05,
    width = 1.45, height = 1.45, just = c("left", "top")
)
plotGG(
    plot = vaccines_FLpie,
    x = 6.37, y = 1.67,
    width = 1.45, height = 1.45, just = c("left", "top")
)
```

We can also easily add additional elements to further enhance our 
complex `ggplot` arrangments, like a precise placement of text labels:

```{r eval = FALSE, echo = TRUE}
plotText(
    label = c("not", "partially", "fully vaccinated"),
    fontfamily = "ProximaNova", fontcolor = "black", fontsize = 7,
    x = c(6.58, 7.3, 7.435),
    y = c(0.74, 1.12, 0.51), just = c("left", "bottom")
)
plotText(
    label = c("not", "partially", "fully vaccinated"),
    fontfamily = "ProximaNova", fontcolor = "black", fontsize = 7,
    x = c(6.58, 7.39, 7.435),
    y = c(2.47, 2.75, 2.2), just = c("left", "bottom")
)
plotText(label = paste("Sources: The COVID Tracking Project;",
                        "Johns Hopkins Center for Civic Impact"),
            fontfamily = "ProximaNova", fontcolor = "black",
            fontsize = 7, fontface = "italic",
            x = 0.15, y = 3.45, just = c("left", "bottom"))
```

We are then left with a complex, precise, and elegant arrangement of 
`ggplots` as if we had arranged them together with graphic design software:

```{r gg_plot07, echo=FALSE, fig.height=4.5, fig.showtext=TRUE, fig.width=11.5}
pageCreate(
    width = 9.5, height = 3.5, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
plotGG(
    plot = US_map,
    x = 0.1, y = 0,
    width = 6.5, height = 3.5, just = c("left", "top")
)
plotGG(
    plot = cases_NYline,
    x = 6.25, y = 1.8,
    width = 3.025, height = 1.4, just = c("left", "bottom")
)
plotGG(
    plot = cases_FLline,
    x = 6.25, y = 3.5,
    width = 3.025, height = 1.4, just = c("left", "bottom")
)
plotGG(
    plot = vaccines_NYpie,
    x = 6.37, y = -0.05,
    width = 1.45, height = 1.45, just = c("left", "top")
)
plotGG(
    plot = vaccines_FLpie,
    x = 6.37, y = 1.67,
    width = 1.45, height = 1.45, just = c("left", "top")
)
plotText(
    label = c("not", "partially", "fully vaccinated"),
    fontfamily = "ProximaNova", fontcolor = "black", fontsize = 7,
    x = c(6.58, 7.3, 7.435),
    y = c(0.74, 1.12, 0.51), just = c("left", "bottom")
)
plotText(
    label = c("not", "partially", "fully vaccinated"),
    fontfamily = "ProximaNova", fontcolor = "black", fontsize = 7,
    x = c(6.58, 7.39, 7.435),
    y = c(2.47, 2.75, 2.2), just = c("left", "bottom")
)
plotText(label = paste("Sources: The COVID Tracking Project;",
                        "Johns Hopkins Center for Civic Impact"),
            fontfamily = "ProximaNova", fontcolor = "black",
            fontsize = 7, fontface = "italic",
            x = 0.15, y = 3.45, just = c("left", "bottom"))
```

# Session Info
```{r}
sessionInfo()
```
