---
title: Ideograms and Shapes
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Ideograms and Shapes}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    message = FALSE,
    warning = FALSE
)
library(BentoBox)
library(grid)
library(BentoBoxData)
```

Beyond providing functions for plotting and arranging various genomic 
datasets, `BentoBox` also gives users the functionality to plot other elements
within a `BentoBox` page layout:

## Ideograms

In addition to a genomic axis label, it can also be useful to include an 
ideogram representation of a chromosome to give a broader context of the 
location of genomic data. UCSC Giemsa stain cytoband information for various 
genomic assemblies is retrieved from `AnnotationHub` for default assembly, 
but users can also provide their own Giemsa stain information if they desire.

Ideograms can be plotted both vertically and horizontally:
```{r ideogram_01, echo=TRUE, fig.height=5.25, fig.width=10.35, message=FALSE}
library(AnnotationHub)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(GenomeInfoDb)

## Get sizes of chromosomes to scale their sizes
tx_db <- TxDb.Hsapiens.UCSC.hg19.knownGene
chromSizes <- GenomeInfoDb::seqlengths(tx_db)
maxChromSize <- max(chromSizes)

bbPageCreate(
    width = 8.35, height = 4.25, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
xCoord <- 0.15
for (chr in c(paste0("chr", seq(1, 22)), "chrX", "chrY")) {
    height <- (4 * chromSizes[[chr]]) / maxChromSize
    bbPlotIdeogram(
        chrom = chr, assembly = "hg19",
        orientation = "v",
        x = xCoord, y = 4,
        width = 0.2, height = height,
        just = "bottom"
    )
    bbPlotText(
        label = gsub("chr", "", chr),
        x = xCoord, y = 4.1, fontsize = 10
    )
    xCoord <- xCoord + 0.35
}
```

```{r ideogram_02, echo=TRUE, fig.height=.5, fig.width=8.25, message=FALSE}
bbPageCreate(
    width = 6.25, height = 0.5, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)

bbPlotIdeogram(
    chrom = "chr1", assembly = "hg19",
    orientation = "h",
    x = 0.25, y = unit(0.25, "npc"), width = 5.75, height = 0.3
)
```

The cytobands can also be hidden if a more minimal ideogram is preferred:

```{r eval=FALSE, echo=TRUE, message=FALSE}
bbPlotIdeogram(
    showBands = FALSE,
    chrom = "chr1", assembly = "hg19",
    orientation = "h",
    x = 0.25, y = unit(0.25, "npc"), width = 5.75, height = 0.3
)
```

```{r ideogram_03, echo=FALSE, fig.height=.5, fig.width=8.25, message=FALSE}
bbPageCreate(
    width = 6.25, height = 0.5, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)

bbPlotIdeogram(
    showBands = FALSE,
    chrom = "chr1", assembly = "hg19",
    orientation = "h",
    x = 0.25, y = unit(0.25, "npc"), width = 5.75, height = 0.3
)
```

To highlight a specific genomic region on an ideogram, see the 
article [An in depth look at plot annotations](annotations.html#genomic-region-highlights-and-zooms-1). 

## Images and basic shapes

`BentoBox` also allows users to plot images and basic shapes and elements to 
further enhance and customize plot layouts. The functions `bbPlotCircle()`, 
`bbPlotPolygon()`, `bbPlotRaster()`, `bbPlotRect()`, `bbPlotSegments()`, 
and `bbPlotText()` provide an intuitive way to plot basic `grid` `grobs` 
without requiring any knowledge of `grid` graphics.

For example, we can include the `BentoBox` mascot Edamaman in our figures!

```{r shapes, echo=TRUE, fig.height=7, fig.showtext=TRUE, fig.width=6}
library(png)
library(showtext)
font_add(
    family = "ProximaNova",
    regular = system.file("extdata",
        "proximanova-regular.otf",
        package = "BentoBox"
    ),
    bold = system.file("extdata",
        "proximanova-semibold.otf",
        package = "BentoBox"
    )
)
showtext_auto()

edamaman <- readPNG(system.file("images",
    "bento-edamaman.png",
    package = "BentoBox"
))
logotype <- readPNG(system.file("images",
    "bento-logotype-singleline-black.png",
    package = "BentoBox"
))

bbPageCreate(
    width = 5, height = 6, default.units = "inches",
    showGuides = FALSE, xgrid = 0, ygrid = 0
)
bbPlotRaster(
    image = logotype,
    x = 2.5, y = 0.25, width = 3.25, height = 0.5, just = "top"
)
bbPlotRaster(
    image = edamaman,
    x = 2.5, y = 5.5, width = 2, height = 4, just = "bottom"
)
bbPlotText(
    label = "Edamaman",
    fontsize = 20, fontfamily = "ProximaNova", fontface = "bold",
    x = 2.5, y = 0.9, just = "top"
)
```

For more detailed usage of basic shape functions, see the function-specific 
reference examples with `?function()` (e.g. `?bbPlotCircle()`).

## Other grob objects

We saw how to add `ggplots` to BentoBox layouts in the 
vignette [Incoporating ggplots](incoporating_ggplots.html) with the
`bbPlotGG()` function. Beyond customizing the coordinates and dimensions of
`ggplots`, `bbPlotGG()` can also be used to incorporate other `grob` and
`gtable` objects. Thus, `BentoBox` allows us to easily mix and arrange most
kinds of plot objects for ultimate customizability.

# Session Info
```{r}
sessionInfo()
```